---
title: "kableExtra-Pkg-NV-LQ"
author: "Maria Guideng"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = FALSE)
```

 -----------------------------------------------------------------
                       kableExtra-Pkg-NV-LQ                       
 -----------------------------------------------------------------
  Script purpose:                                                 
  Format and render a table using `knitr::kable` and `kableExtra` 
 -----------------------------------------------------------------

LQ calculations were performed in the previous script (filename: _2script-lq-model.R_). This final script is specifically for presentation purposes and not for further analysis.


### Prelims
##################################################################
```{r, results='hide', message=FALSE, warning=FALSE}
library(knitr)
options(knitr.table.format = "html")

library(kableExtra)
ls("package:kableExtra")
#help(package = "kableExtra")

library(magrittr)     # %>%
library(dplyr)        # to mutate
library(formattable)  # to add gradient to table cell specs
```


### Prepare Data Table
##################################################################

Bring in tabular data of LQ table
```{r}
urlfile <- 'https://raw.githubusercontent.com/mguideng/KableExtra-Pkg-NV-LQ/master/data/lqmod.csv'
lqtab <- read.csv(urlfile)
```

Change class modes
```{r}
sapply(lqtab, class)

cols.num <- c("US.Empl", "US.Percent", "LV.Empl", "LV.Percent", "LV.EmplRqmts", "LV.ExcessEmpl")
lqtab[cols.num] <- sapply(lqtab[cols.num],as.numeric)

cols.char <- c("NAICS", "Industry", "LV.LQ", "LV.LQClass")
lqtab[cols.char] <- sapply(lqtab[cols.char],as.character)
```

Add a row for totals
```{r results='hide', message=FALSE, warning=FALSE}
library(janitor)
lqtab <- lqtab %>% 
  adorn_totals("row", fill="", na.rm = F)

# Revert back to a numeric. Was set as characters because did not want summed in adorn_totals()
lqtab$LV.LQ <- as.numeric(lqtab$LV.LQ)
```

Format to include percents, commas and decimals for clarity
```{r}
# Percent columns
lqtab$US.Percent <- paste(round(lqtab$US.Percent*100, digits=2),"%",sep="")
lqtab$LV.Percent <- paste(round(lqtab$LV.Percent*100, digits=2),"%",sep="")

# Number columns
lqtab$US.Empl <- format(lqtab$US.Empl,big.mark=",", trim=TRUE)
lqtab$LV.Empl <- format(lqtab$LV.Empl,big.mark=",", trim=TRUE)
lqtab$LV.EmplRqmts <- format(lqtab$LV.EmplRqmts,big.mark=",", trim=TRUE)
lqtab$LV.ExcessEmpl <- format(lqtab$LV.ExcessEmpl,big.mark=",", trim=TRUE)

# Quotient column
lqtab$LV.LQ <- format(round(lqtab$LV.LQ, 2), nsmall = 2)
```

Name headers that will include spaces in them for clarity
```{r}
names(lqtab)
colnames(lqtab) = c("NAICS", "Industry", "US Employment", "US % Share", "Las Vegas Employment",
                   "Las Vegas % Share", "Employment for Las Vegas Requirements", "Las Vegas LQ",
                   "Las Vegas LQ Classification", "Las Vegas Excess Employment (to export) or Deficit")
```


### Rendering kableExtra Formatted Tables
##################################################################

This is the basic HTML table in `kable`:

```{r}
kable(lqtab, "html")
```

And here's the makeover after some formatting help from `kableExtra`.
(Since this will be duplicated for different years, change year in the title caption accordingly, i.e., "20XX Las Vegas MSA Location Quotient".)

```{r, warning=F}
lqtab %>% 
  mutate(
    `Industry` = cell_spec(`Industry`, color = ifelse(`Industry` %in% c("RE", "Management", "Admin", "Arts", "Accommodation"),"black","")),
    `Las Vegas LQ` = color_tile("white","#00B2FF")(`Las Vegas LQ`),
    `Las Vegas LQ Classification` = cell_spec(`Las Vegas LQ Classification`, bold=T, color="black", background = ifelse(`Las Vegas LQ Classification`== "Export","#81D4FA","")), 
    `Las Vegas Excess Employment (to export) or Deficit` = cell_spec(`Las Vegas Excess Employment (to export) or Deficit`, color=ifelse(`Las Vegas Excess Employment (to export) or Deficit` > 1,"","red"))

) %>%
  select(NAICS, Industry, `US Employment`, `US % Share`, `Las Vegas Employment`, `Las Vegas % Share`,
         `Employment for Las Vegas Requirements`, `Las Vegas LQ`, `Las Vegas LQ Classification`,  
         `Las Vegas Excess Employment (to export) or Deficit`) %>%
  kable(escape = F, align = "c", caption = "2016 Las Vegas MSA Location Quotient", linesep = "") %>%
  kable_styling(bootstrap_options = c("hover", "condensed"), 
                position = "center", font_size = 11, full_width = T) %>% 
  column_spec(1, width = "1cm", border_left = T) %>% 
  column_spec(2, width = "2cm", border_right = T) %>% 
  column_spec(3, width = "2cm") %>% 
  column_spec(4, width = "1cm", border_right = T) %>% 
  column_spec(5, width = "2cm") %>% 
  column_spec(6, width = "1cm", border_right = T) %>% 
  column_spec(7, width = "2cm", border_right = T) %>%
  column_spec(8, width = "2cm", border_right = T, bold = T, color = "black") %>%
  column_spec(9, width = "2cm") %>% 
  column_spec(10, width = "2cm", border_right = T) %>% 
  row_spec(0, bold = T) %>% 
  row_spec(11, bold = T, background = "#F2F4F4") %>% 
  row_spec(13:14, bold = T, background = "#F2F4F4") %>% 
  row_spec(17:18, bold = T, background = "#F2F4F4") %>% 
  row_spec(21, bold = T) %>% 
  footnote(general = c("Assumes Las Vegas demands goods/services at the same level as the nation.  Totals may not sum exactly due to roundng."), general_title = "Notes: ")
```

